### 2011-12-05 abs.tex
Went all Unicode, all the time.

This dissertation considers two types of ``gaps'' in the lexicon. In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of gapped structures. The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English. A number of prior studies have argued that the grammar must have a separate ``phonotactic'' module to account

at the level of prosodic word. In both cases, it is argued that the i

>>>>

This dissertation considers two types of gaps in the lexicon. In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of missing structures.

The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English. Many prior studies have argued for morpheme structure constraints, generalizations that account for the underlying shape of morphs. It is shown, however, that many of the gaps in the inventory of medial consonant clusters are due to sampling from sparse data, cannot be reliably predicted by current computational models of phonotactic learning, and can thus be regarded as accidental gaps. In contrast, phonological alternations induce reliable structural gaps in the inventory of medial clusters. This suggests that language-specific phonotactics can be derived without a separate phonotactic module (\citealt{Halle1962}). 
%This model is validated with a meta-analysis of 5 English wordlikeness studies.

The second part considers the inflectional patterns in a number of languages. There are many languages in which a certain class of words unexpectedly fails to 
Hungarian and Kinande.

The use of elsewhere rules to account for allomorphy fail to

This model is validated using data on inflectional gaps in English, Spanish, Russian, Swedish, and Greek.

% Keywords: Gaps, sparsity, Zipf's Law, acquisition, phonotactics, wordlikeness, syllable contact, inflectional gaps, defectivity, lexical exceptions
% Languages: Turkish, English, Latin, Hungarian, Kinande, Spanish, Russian, Swedish, Greek

### 2011-12-05 ack.tex
Went all Unicode, all the time.

I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. Elsewhere, I have been shown a great deal of kindness by so many, but especially Morgan Sonderregger and Maryam Bakht. Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. Among non-linguists, my good friend Carly Fisher stands out for her unflagging support. Among parents, mine stand out for having more faith in their son than he could ever merit.

>>>>

I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. Elsewhere, I have been shown a great deal of kindness by so many, but especially Maryam Bakht. Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. Katherine McEldoon provided a useful review of numerical cognition research. Carly Fisher gave me her unflagging support.

### 2011-12-05 clusters.tex
Went all Unicode, all the time.

@@

@@

\ex Artificial language repetitions \citep[from][]{Wright1975}: \\
\begin{tabular}{l l l l l l} %\toprule
%a. & [go\textupsilon np] & $>$ & [gu\lm mp] \\
%[g\textturnv np]  
%b. & [\dh \textturnv\textipa{N}d] & $>$ & [\dh \textopeno\textipa{N}g\textschwa] \\ % $>$ [t\textopeno\textipa{N}g\textschwa] 
%c. & [\textdyoghlig umg] & $>$ & [\textdyoghlig u\textipa{N}g\textschwa] \\
\end{tabular} \xe

@@

t
If this task is viewed as something like loanword adaptation, then there are certainly good reasons to think that extragrammatical pressures \citep{Dupoux1999,Ussishkin2003,Peperkamp2008} could be responsible for what is observed. That said, it is certainly noteable that

@@

A major result of the American structuralists was the finding that generalizations which hold of the sound sequences found in underlying forms may not hold across morpheme boundaries. If this is correct, then there must be some sense in which phonology is sensitive to the presence of morpheme boundaries. In his history of phonology, \citet[][267]{Anderson1985} credits this innovation to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [\c{c}] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm x\textschwa n] `cake' $\sim$ \emph{Kuhchen} [ku\lm\c{c}\textschwa n] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [\c{c}] and [x] can be cast as allophones of the same phoneme (likely /\c{c}/, the basic variant).

@@

\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\ 
\begin{tabular}{r l l l} %\toprule
   & MSC & complex exception \\ %\midrule
%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
%b. & *{C}\textipa{@}{C}e & [n\`i-k\`\textschwa b\textschwa-de] & `who entered'        \\
%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
%d. & *{C}i{C}e & [te-n\`i-kee-t\`\textschwa] & `and went away'      \\ %\toprule
a. & *{C}a{C}e & [k\'a\textsuperscript{n}dee] & `kept \ldots inside' \\
b. & *{C}\textipa{@}{C}e & [n\`ik\`\textschwa b\textschwa de] & `who entered'        \\
c. & *{C}e{C}i & [ten\'i ke\textsuperscript{n}da] & `was walking         \\
d. & *{C}i{C}e & [ten\`i keet\`\textschwa] & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
\end{tabular} \xe

@@

While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later. \citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. For instance, most Russian infinitives end with a final /t'/. The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\textceltpal -s\ldots/ juncture triggers the loss of palatalization. However, palatalization of a stem-final labial or coronal---including /t'/, as in (\nextx c)---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses a highly abstract phonemic transcription. The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers. Palatalization is marked even where not constrastive.}

\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
\begin{tabular}{l l l l} %\toprule
   &  infinitive & imperative \\ %\midrule
a. & [slav\textceltpal its\textschwa]         & [slaf\textceltpal s\textschwa]            & `be glorious'    \\
   & [pramits\textschwa]         & [upram\textceltpal s\textschwa]          & `be stubborn'    \\
b. & [kras\textceltpal its\textschwa]         & [kras\textceltpal s\textschwa]           & `put on makeup'  \\
   & [\textyogh arits\textschwa]       & [\textyogh ar\textceltpal s\textschwa]         & `roast'          \\ 
c. & [z\textschwa byts\textschwa] & [z\textschwa but\textceltpal s\textschwa] & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular} \xe

@@

Both forms in (\lastx c) contain a /t\textceltpal-s/ juncture, though only one is faithful. \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization. A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{K\"uche} `kitchen; cuisine'). There are obvious parallels between \citeauthor{Jakobson1932}'s analysis and notions like readjustment in \emph{SPE}, or cyclic rules in lexical phonology or the \citet{Halle1987} model.

@@

\ex Feature specification for Turkish vowels: \par\nobreak
\begin{tabular}{l c c c c}
 & \multicolumn{2}{c}{[$-$\textsc{back}]} & \multicolumn{2}{c}{[$+$\textsc{back}]} \\
& [$-$\textsc{round}] & [$+$\textsc{round}] & [$-$\textsc{round}] & [$+$\textsc{round}] \\
\buf[$+$\textsc{high}] & \emph{i} [i]        & \emph{\"u} [y]      & \emph{\i} [\textturnm] & \emph{u} [u] \\
\buf[$-$\textsc{high}] & \emph{e} [e]        & \emph{\"o} [\o]     & \emph{a} [\textschwa]  & \emph{o} [o] \\
\end{tabular}
\xe

%into four [$+$\textsc{back}] vowels \emph{u}, \emph{\i}, \emph{o}, \emph{a}, denoting [u, \textturnm, o, a] and four [$-$\textsc{back}] vowels \emph{i}, \emph{\"u}, \emph{e}, \emph{\"o} denoting [i, y, e, \o].

Native roots generally contain only [$+$\textsc{back}] vowels (e.g., \emph{g\"orev} `duty') or [$-$\textsc{back}] vowels (e.g., \emph{odun} `wood'), but loanwords need not conform to this generalization (e.g., \emph{sosis} `sausage'). Despite this latter fact, there is some evidence that this constraint on roots is an induced one.

\ex Turkish declensional suffix allomorphy (after \citealt[][315]{Zimmer1969}): \par\nobreak
\begin{tabular}{l l l l l}
   & nom.sg.      & acc.sg.         & gen.sg.              \\
a. & \emph{it}    & \emph{iti}      & \emph{itin}      & `dog'  \\
%a. & \emph{g\"ul} & \emph{g\"ul\"u} & \emph{g\"ul\"un} & `rose' \\
c. & \emph{k\i z} & \emph{k\i z\i}  & \emph{k\i z\i n} & `girl' \\
%c. & \emph{tuz}   & \emph{tuzu}     & \emph{tuzun}     & `salt' \\
\end{tabular}
\xe

\noindent
Similar facts are indicated by the rest of the declensional paradigms. It is possible to analyze the forms of the acc.sg. 
(\emph{i}, \emph{\i})
%(\emph{\"u}, \emph{u})
%(\emph{i}, \emph{\"u}, \emph{\i}, \emph{u})
and gen.sg. 
(\emph{in}, \emph{\i n})
%(\emph{\"un}, \emph{un})
%(\emph{in}, \emph{\"un}, \emph{\i n}, and \emph{un}) 
without suppletion, using the following phonological rule:

@@

In other words, a vowel takes on the [\textsc{back}] specifications of the preceding vowel. The acc.sg. suffix is then [$+$\textsc{high}] (and for these cases, [$-$\textsc{round}]); there is no evidence bearing on whether it is one of /i, \i/ or simply underspecified for [back]. The gen.sg. suffix is the same vowel followed by /n/.

@@

Another morpheme structure constraint discussed by \citeauthor{Zimmer1969} is that a vowel immediately preceded by \emph{a} and a labial consonant (\emph{p}, \emph{b}, \emph{f}, \emph{v}) must be [$+$\textsc{round}], though once again there are a few roots that are exceptions to this generalization. What is crucial here is that this labial consonant constraint is not involved in any suffix alternation, and in fact, does not hold of complex words. For instance, the acc.sg. of \emph{av} `hunt' is not *\emph{avu}, but rather \emph{av\i}. This suggests that this constraint cannot be considered an induced restriction.

@@

%In Zimmer's first experiment, native Turkish speakers were given a questionnaire where each line consisted of two nonce words written in the modern orthography. The subjects were asked to read the words allowed and to put a checkmark next to the word they preferred. Subjects were also permitted to check both words, indicating ``no preference''. Four of the pairs in the first study consisted of a pair of nonce words differing only whether or not they are consistent with the [\textsc{back}] \textsc{Harmony} generalization. These results are given in Table \ref{backharm}.
%\begin{table}
%\centering
%\begin{tabular}{l r l r r}
%\toprule
%hi \\
%\midrule
%\emph{temez}   & $19$ & \emph{temaz} & $3$ & $1$ \\ %back
%%\emph{tipez}   & $21$ & \emph{tipaz} & $1$ & $1$ \\ %back
%\emph{ter\"uz} & $20$ & \emph{teruz} & $1$ & $2$ \\ %back
%%t\"or\"uz & $19$ & t\"oriz & $1$ & $3$ \\ %round
%\emph{teriz}   & $23$ & ter\i z & $0$ & $0$ \\ 
%% tamuz   & $3$  & tam\I z & $16$ \\
%% tafuz   & $3$  & taf\I z & $17$ \\
%% tapuz   & $7$  & tap\I z & $9$ \\
%\bottomrule
%\end{tabular}
%\caption{This is a test, \citep[][313]{Zimmer1969}}
%\label{backharm}
%\end{table}

%\begin{tabular}{l r l r r}
%\toprule
%hi \\
%\midrule
%tamuz & $3$ & tam\i z & $16$ & $4$ \\
%tafuz & $3$ & taf\i z & $17$ & $3$ \\
%tapuz & $7$ & tap\i z & $9$  & $7$ \\
%tavuz & $9$ & tav\i z & $4$  & $10$ \\
%tabuz & $5$ & tab\i z & $12$ & $6$ \\
%\bottomrule
%\end{tabular}
%\caption{\citet[][313]{Zimmer1969}}
%\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l r l r r}
\toprule
harmonic      & \#   & disharmonic & \#   \\
\midrule
\emph{pemez}  & $30$  & \emph{pemaz}     & $2$  \\
\emph{teriz}   & $28$ & \emph{ter\i z}   & $3$  \\
\emph{tokaz}   & $26$ & \emph{tokez}     & $6$  \\
\emph{tipez}   & $24$ & \emph{tipaz}     & $8$  \\
\emph{ter\"uz} & $19$ & \emph{teruz}     & $13$ \\
\bottomrule
\end{tabular}
\caption{this is at test \citet[][314]{Zimmer1969}}
\label{backharm}
\end{table}
% c = 127, d = 32, p = 5.233e-29

@@

\begin{table}
\centering
\begin{tabular}{l r l r}
\toprule
harmonic & \# & disharmonic & \# \\
\midrule
\emph{tafuz} & $21$ & \emph{taf\i z} & $11$ \\
\emph{pamuz} & $17$ & \emph{pam\i z} & $15$ \\
\emph{tapuz} & $17$ & \emph{tap\i z} & $15$ \\
\emph{mavuz} & $16$ & \emph{mav\i z} & $16$ \\
\emph{tabuz} & $16$ & \emph{tab\i z} & $16$ \\
\bottomrule
\end{tabular}
\caption{this is at test \citet[][314]{Zimmer1969}}
\label{labcons}
\end{table}
% c = 87, d = 73, p = 0.101

Here is it clear that the preferences, while going in the expected direction, are far smaller ($\gamma = 0.088$). There are ties for two of the lexical items, and only one of the five pairs, \emph{tafuz} $\sim$ \emph{taf\i z}, matches the stronger preferences seen with backness harmony. If we remove this pair, $\gamma = 0.032$.  It might be possible to though any failure to find significance here should surely be attributed to insufficient statistical power, not evidence for the null hypothesis.

@@

>>>>

\citet[][111]{Vennemann1972}
\citet[][94f.]{Shibatani1973}
\citet[][84]{Hooper1973}
\citet{Kaye1974}
\citet{Clayton1976}

Lang Acq

\citet{White2008}

%\citeauthor{Hooper1973} (ibid., p.~84) provides an example from the analysis of nasal place assimilation in Spanish given by \citet[][16]{Harris1969}

Odawa provides

peripherality
\citet[][8f.]{LGB}
\citep{Duanmu2008}
\citet{Borowsky1989}

\section{Evidence in phonotactics}

\subsection{Evidence from acquisition}

@@

\subsection{Evidence from wordlikeness judgements}

\subsection{Evidence from lexical statistics}

\citet{Berkley1994a,Berkley1994b}

\begin{quotation}
\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][]{Fischer-Jorgensen1952}
\end{quotation}

Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters two years later.

\begin{quotation}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots~partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quotation}

\begin{quotation}
\ldots~in languages that allow relatively complex consonant clusters, the occurring clusters represent only a small percentage of all theoretically possible consonant clusters. \citep[][94]{Hooper1973}
\end{quotation}

Similar observations have also been made in recent years, for instance in the study of consonant co-occurrence restrictions in Arabic performed by \citeauthor{Frisch2004}.

\begin{quotation}
In describing and explaining the phonotatic patterns in any language, phonologists recognize a distinctino between accidental and systematic gaps in the lexcion. In practice, this distinction has often been applied in an intuitive and \emph{post-hoc} fashion. A gap is taken to be systematic if it belongs to a natural class of examples according to the current theory of the investigator. Otherwise, it is viewed as accidental. \citep[][180]{Frisch2004}
\end{quotation}

There are also a number of prior studies \citet{Kaeding1898,Weiss1961,Sigurd1968,Tambovtsev2007} which note that phonemes and phoneme sequences are sparsely distributed: that is, there are a few very frequent phonemes or sequences, and many infrequent phonemes or sequences.

@@

\ex Artificial language repetitions \citep[from][]{Wright1975}: \\
\begin{tabular}{l l l l l l} %\toprule
%a. & [go\textupsilon np] & $>$ & [gu\lm mp] \\
%[g\textturnv np]  
%b. & [\dh \textturnv\textipa{N}d] & $>$ & [\dh \textopeno\textipa{N}gə] \\ % $>$ [t\textopeno\textipa{N}gə] 
%c. & [\textdyoghlig umg] & $>$ & [\textdyoghlig u\textipa{N}gə] \\
\end{tabular} \xe

@@

t
If this task is viewed as something like loanword adaptation, then there are certainly good reasons to think that extragrammatical pressures \citep{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008} could be responsible for what is observed. That said, it is certainly noteable that

@@

A major result of the American structuralists was the finding that generalizations which hold of the sound sequences found in underlying forms may not hold across morpheme boundaries. If this is correct, then there must be some sense in which phonology is sensitive to the presence of morpheme boundaries. In his history of phonology, \citet[][267]{Anderson1985} credits this innovation to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm xən] `cake' $\sim$ \emph{Kuhchen} [ku\lm çən] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [ç] and [x] can be cast as allophones of the same phoneme (likely /ç/, the basic variant).

@@

\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\ 
\begin{tabular}{r l l l} %\toprule
   & MSC & complex exception \\ %\midrule
%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
\end{tabular} \xe

@@

While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later. \citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. For instance, most Russian infinitives end with a final /t\pal/. The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\pal-s\ldots/ juncture triggers the loss of palatalization. However, palatalization of a stem-final labial or coronal---including /t'/, as in (\nextx c)---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses a highly abstract phonemic transcription. The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers. Palatalization is marked even where not constrastive.}

\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
\begin{tabular}{l l l l} %\toprule
   &  infinitive & imperative \\ %\midrule
a. & [slav\pal itsə] & [slaf\pal sə] & `be glorious'    \\
   & [pramitsə]              & [upram\pal sə]          & `be stubborn'    \\
b. & [kras\pal itsə]         & [kras\pal sə]           & `put on makeup'  \\
   & [ʒaritsə]       & [ʒar\pal sə]         & `roast'          \\ 
c. & [zəbytsə] & [zəbut\pal sə] & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular} \xe

@@

Both forms in (\lastx c) contain a /t\pal-s/ juncture, though only one is faithful. \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization. A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{Küche} `kitchen; cuisine'). There are obvious parallels between \citeauthor{Jakobson1932}'s analysis and notions like readjustment in \emph{SPE}, or cyclic rules in lexical phonology or the \citet{Halle1987} model.

@@

\ex Feature specification for Turkish vowels: \par\nobreak
\begin{tabular}{l c c c c}
 & \multicolumn{2}{c}{[$-$\textsc{back}]} & \multicolumn{2}{c}{[$+$\textsc{back}]} \\
& [$-$\textsc{round}] & [$+$\textsc{round}] & [$-$\textsc{round}] & [$+$\textsc{round}] \\
\buf[$+$\textsc{high}] & \emph{i} [i]        & \emph{ü} [y]      & \emph{ı} [ɯ] & \emph{u} [u] \\
\buf[$-$\textsc{high}] & \emph{e} [e]        & \emph{ö} [ø]     & \emph{a} [ə]  & \emph{o} [o] \\
\end{tabular}
\xe

%into four [$+$\textsc{back}] vowels \emph{u}, \emph{ı}, \emph{o, \emph{a}, denoting [u, ɯ, o, a] and four [$-$\textsc{back}] vowels \emph{i}, \emph{ü}, \emph{e}, \emph{ö} denoting [i, y, e, ø].

Native roots generally contain only [$+$\textsc{back}] vowels (e.g., \emph{görev} `duty') or [$-$\textsc{back}] vowels (e.g., \emph{odun} `wood'), but loanwords need not conform to this generalization (e.g., \emph{sosis} `sausage'). Despite this latter fact, there is some evidence that this constraint on roots is an induced one.

\ex Turkish declensional suffix allomorphy (after \citealt[][315]{Zimmer1969}): \par\nobreak
\begin{tabular}{l l l l l}
   & nom.sg.      & acc.sg.         & gen.sg.              \\
a. & \emph{it}    & \emph{iti}      & \emph{itin}      & `dog'  \\
%a. & \emph{gül} & \emph{gülü} & \emph{gülün} & `rose' \\
c. & \emph{kız} & \emph{kızı}  & \emph{kızın} & `girl' \\
%c. & \emph{tuz}   & \emph{tuzu}     & \emph{tuzun}     & `salt' \\
\end{tabular}
\xe

\noindent
Similar facts are indicated by the rest of the declensional paradigms. It is possible to analyze the forms of the acc.sg. 
(\emph{i}, \emph{ı})
%(\emph{ü}, \emph{u})
%(\emph{i}, \emph{ü}, \emph{ı}, \emph{u})
and gen.sg. 
(\emph{in}, \emph{ı n})
%(\emph{ün}, \emph{un})
%(\emph{in}, \emph{ün}, \emph{ı n}, and \emph{un}) 
without suppletion, using the following phonological rule:

@@

In other words, a vowel takes on the [\textsc{back}] specifications of the preceding vowel. The acc.sg. suffix is then [$+$\textsc{high}] (and for these cases, [$-$\textsc{round}]); there is no evidence bearing on whether it is one of /i, ı/ or simply underspecified for [back]. The gen.sg. suffix is the same vowel followed by /n/.

@@

Another morpheme structure constraint discussed by \citeauthor{Zimmer1969} is that a vowel immediately preceded by \emph{a} and a labial consonant (\emph{p}, \emph{b}, \emph{f}, \emph{v}) must be [$+$\textsc{round}], though once again there are a few roots that are exceptions to this generalization. What is crucial here is that this labial consonant constraint is not involved in any suffix alternation, and in fact, does not hold of complex words. For instance, the acc.sg. of \emph{av} `hunt' is not *\emph{avu}, but rather \emph{avı}. This suggests that this constraint cannot be considered an induced restriction.

@@

%In Zimmer's first experiment, native Turkish speakers were given a questionnaire where each line consisted of two nonce words written in the modern orthography. The subjects were asked to read the words allowed and to put a checkmark next to the word they preferred. Subjects were also permitted to check both words, indicating ``no preference''. Four of the pairs in the first study consisted of a pair of nonce words differing only whether or not they are consistent with the [\textsc{back}] \textsc{Harmony} generalization. These results are given in Table \ref{backharm}.
%\begin{table}
%\centering
%\begin{tabular}{l r l r r}
%\toprule
%hi \\
%\midrule
%\emph{temez}   & $19$ & \emph{temaz} & $3$ & $1$ \\ %back
%%\emph{tipez}   & $21$ & \emph{tipaz} & $1$ & $1$ \\ %back
%\emph{terüz} & $20$ & \emph{teruz} & $1$ & $2$ \\ %back
%%törüz & $19$ & töriz & $1$ & $3$ \\ %round
%\emph{teriz}   & $23$ & terı z & $0$ & $0$ \\ 
%% tamuz   & $3$  & tam\I z & $16$ \\
%% tafuz   & $3$  & taf\I z & $17$ \\
%% tapuz   & $7$  & tap\I z & $9$ \\
%\bottomrule
%\end{tabular}
%\caption{This is a test, \citep[][313]{Zimmer1969}}
%\label{backharm}
%\end{table}

%\begin{tabular}{l r l r r}
%\toprule
%hi \\
%\midrule
%tamuz & $3$ & tamı z & $16$ & $4$ \\
%tafuz & $3$ & tafı z & $17$ & $3$ \\
%tapuz & $7$ & tapı z & $9$  & $7$ \\
%tavuz & $9$ & tavı z & $4$  & $10$ \\
%tabuz & $5$ & tabı z & $12$ & $6$ \\
%\bottomrule
%\end{tabular}
%\caption{\citet[][313]{Zimmer1969}}
%\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l r l r r}
\toprule
harmonic      & \#   & disharmonic   & \#   \\
\midrule
\emph{pemez} & $30$ & \emph{pemaz} & $2$  \\
\emph{teriz} & $28$ & \emph{terız} & $3$  \\
\emph{teriz} & $28$ & \emph{terız} & $3$  \\
\emph{tokaz} & $26$ & \emph{tokez} & $6$  \\
\emph{tipez} & $24$ & \emph{tipaz} & $8$  \\
\emph{terüz} & $19$ & \emph{teruz} & $13$ \\
\bottomrule
\end{tabular}
\caption{this is at test \citet[][314]{Zimmer1969}}
\label{backharm}
\end{table}
% c = 127, d = 32, p = 5.233e-29

@@

\begin{table}
\centering
\begin{tabular}{l r l r}
\toprule
harmonic & \# & disharmonic & \# \\
\midrule
\emph{tafuz} & $21$ & \emph{tafız} & $11$ \\
\emph{pamuz} & $17$ & \emph{pamız} & $15$ \\
\emph{tapuz} & $17$ & \emph{tapız} & $15$ \\
\emph{mavuz} & $16$ & \emph{mavız} & $16$ \\
\emph{tabuz} & $16$ & \emph{tabız} & $16$ \\
\bottomrule
\end{tabular}
\caption{this is at test \citet[][314]{Zimmer1969}}
\label{labcons}
\end{table}
% c = 87, d = 73, p = 0.101

Here is it clear that the preferences, while going in the expected direction, are far smaller ($\gamma = 0.088$). There are ties for two of the lexical items, and only one of the five pairs, \emph{tafuz} $\sim$ \emph{tafı z}, matches the stronger preferences seen with backness harmony. If we remove this pair, $\gamma = 0.032$.  It might be possible to though any failure to find significance here should surely be attributed to insufficient statistical power, not evidence for the null hypothesis.

@@

soft-margin support vector machine \citep{Cortes1995}

### 2011-12-05 lexinsertion.tex
Went all Unicode, all the time.

\citep{Pater2000,Pater2005b,Pater2005a,Pater2006,Pater2009,Becker2009a}

>>>>

\section{The reality of inflectional gaps}

Comparing the counts of the preterite and non-preterite forms of \emph{undergo} and \emph{forgo} in both corpora with the Fisher exact test demonstrates that the non-usage of a preterite of \emph{forgo} is unlikely to be accidental (BNC: $p = 3.4$\textsc{e}-31, COCA: $p = 1.7$\textsc{e}-186). This is even true of \emph{sightsee}, which occurs with low frequency as a verb (BNC: $p = 0.0227$).

Surveys of inflectional gaps can be fonud in

as surveyed in \citealt{Fanselow2002}, \citealt{Rice2009b}, and \citealt{Baerman2010b}.

A characteristic of the traditional split between derivational morphology and inflectional morphology is

As \citet{Marantz1997b} notes, there are many cases of productive category-changing morphology. In English the suffix \emph{-ness} productively forms abstract adjectives from nouns, and the suffix \emph{-ōs-} plays a similar productive role in Latin:

\ex Classical Latin denominal adjectives:
\begin{tabular}{l l l l l}
   & nom.sg. noun  &            & nom.sg. masc. adjective & \\
a. & \emph{macula} & `spot'     & \emph{maculōsus}       & `spotted'     \\
b. & \emph{uentus} & `windy'    & \emph{uentōsus}        & `windy'       \\
%c. & \emph{mo:ns}  & `mountain' & \emph{monto:sus}        & `mountainous' \\ 
c. & \emph{lutum } & `mud'      & \emph{lutōsus}         & `muddy'       \\
d. & \emph{ulcus}  & `ulcer'    & \emph{ulcerōsus}      & `uclerous'     \\
\end{tabular} \xe

%\ex English simple past suffixes:
%\begin{tabular}{l l l l l l }
%T[\textsc{Past}] & $\rightleftarrow$ & -\zr & \gap & \{$\surd$\textsc{hit}, $\surd$\textsc{sing}, \ldots\}
%                 & $\rightleftarrow$ & -t   & \gap & \{$\surd$\textsc{leave}, $\surd$\textsc{bend}, \ldots\}
%                 & $\rightleftarrow$ & -d   
%\end{tabular} \xe

Table \ref{fgo} provides some token counts from the 100 million word British National Corpus \citep{BNC,BYUBNC} and the 421 million word Corpus of Contemporary American English \citep{COCA}.

\begin{table}
\centering
\begin{tabular}{l@{} l r r l@{} l r r} 
\toprule
    &            & BNC   & COCA   &   &             & BNC & COCA  \\ 
\midrule
    & undergo    & 613   & 2,757  &   & forgo       & 203 & 1,213 \\
    & undergoes  & 123   & 594    &   & forgoes     & 4   & 89    \\
    & undergoing & 581   & 2,464  &   & forgoing    & 24  & 232   \\
    & underwent  & 550   & 2,440  & * & forwent     & 0   & 0     \\
$*$ & undergoed  & 0     & 0      & * & forgoed     & 0   & 0     \\
    & undergone  & 565   & 1,996  &   & forgone     & 64  & 190   \\ 
\midrule
    & foresee    & 271   & 1,020  &   & sightsee    & 4   & 55    \\
    & foresees   & 58    & 300    &   & sightsees   & 0   & 0     \\
    & foreseeing & 38    & 89     &   & sightseeing & 20  & 5     \\
    & foresaw    & 132   & 431    & * & sightsaw    & 0   & 0     \\
$*$ & foreseed   & 0     & 0      & * & sightseed   & 0   & 0     \\
    & foreseen   & 275   & 568    & * & sightseen   & 0   & 0     \\ 
\midrule
    & ride       & 1,607 & 10,904 &   & stride      & 102 & 402   \\
    & rides      & 230   & 2,495  &   & strides     & 54  & 682   \\
    & riding     & 1,696 & 10,994 &   & striding    & 253 & 715   \\
    & rode       & 1,079 & 5,997  &   & strode      & 641 & 1,752 \\
$*$ & rided      & 1     & 0      & * & strided     & 0   & 2     \\
    & ridden     & 456   & 1,303  & * & stridden    & 1   & 2     \\ 
\bottomrule
\end{tabular}
\caption{Token counts the BNC and COCA for English verbs with inflectional gaps}
\label{fgo}
\end{table}

\citep{Pater2000,Pater2005b,Pater2005a,Pater2006,Pater2009,Becker2009b}
